Electricity has become the most important source of energy in the world. Every year demand for electricity has been increasing with the rise of new digital and production based industies. With the surge in demand, problems regarding the supply of electricity with our current infrastructure especially in developing countrys has become prevelent. Some of the problems in trasmission include harmonic distortion, reactive power control, active power control, voltage fluctuation, voltage sag, voltage swell,and optimal power flow\cite{sharma2023comprehensive}. In order to solve these problems we need FACTS (Flexible AC Transmission Systems) devices. STATCOM (Static Synchronous Compensator) is a shunt connected FACTS device that can be used to solve the above mentioned problems.\\
STATCOM depends upon many variables to achieve optimal performance. Those variables are Optimal Size and Placement, control strategy, harmonics distortion, cooridination with existing devices and enviromental factors. Among these variables optimal size and placement and control strategy are the most important variables which determines the speed and smoothness of the response of the STATCOM. Finding the best values to these variables require a lot of time and computational power.\\
Sanepa Feeder is a distribution feeder in Nepal which is facing many of these problems. This paper focuses on implementing a STATCOM in Sanepa Feeder and find the difference in the performance of the system with different locaitions of implementation and different control strategies.Different control strategy used are GA (Genetic Algorithm) and PSO (Particle Swarm Optimization). This paper aims to provide the insights regarding the impacts of different control strategies and different locations of implementation of STATCOM in the distribution system. The results of this paper can be used by the power system engineers to find the optimal size and placement of STATCOM in the distribution system.\\